---
title: "Brownlows Bets"
author: "Robert Solis"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 


```{r}
BrLw <-function(numClass=35,odds=2,countoDo=637){

days=1:365
count=countoDo
repeatn=numClass
varstor=numeric()
  brownlow = 0
  for (i in 1:countoDo) {
    sam=sample(days,size=numClass,replace = T)
    y=duplicated(sam)
    if (sum(y)>0) brownlow=brownlow+1
    else brownlow = brownlow - odds
  }
 return(brownlow)
}

```





# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->

```{r}
brownlow <- array(100)
for (i in 1:100) {
  brownlow[i] <- BrLw()
}

```

```{r}
sorted_brownlow <- sort(brownlow)
plot(sorted_brownlow, type = "l", xlab = "Index", ylab = "Value",
     main = "Line Plot of Brownlow Numerical Vector")
```
```{r}
favstats(sorted_brownlow)
```
the average earning of brownlow bets are 285.31. 95% of the time. this 95% confidence interval is 2 standard deviations below the mean and 2 standard deviations above the mean. The amount is about 210 to 350 amount of cash.


# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 
```{r}
b_low <- array(100)
for (i in 1:100) {
  b_low[i] <- BrLw(odds=4.36)
}

```
```{r}
sorted_b_low <- sort(b_low)
plot(sorted_b_low, type = "l", xlab = "Index", ylab = "Value",
     main = "Line Plot of Brownlow Numerical Vector")
```

```{r}
favstats(sorted_b_low)
```

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->
the means finding the odds at which the amount he wins on average is equal to the amount he loses, resulting in zero overall profit.

In the provided R code, the BrLw function is modified to take an additional argument odds, which represents the odds offered by Professor Brownlow for the bet. The function then calculates the expected profit based on the given odds.




# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
fair_low <- array(100)
for (i in 1:100) {
  fair_low[i] <- BrLw(numClass = 28)
}

```
```{r}
sorted_fair_low <- sort(fair_low)
plot(sorted_fair_low, type = "l", xlab = "Index", ylab = "Value",
     main = "Line Plot of Brownlow Numerical Vector")
```


```{r}
favstats(sorted_fair_low)
```
The mean of is the average expected profit across all the simulations, indicating that, on average, Professor Brownlow tends to gain  money with this 2:1 bet. The standard deviation of measures the variability or dispersion of the expected profits around the mean. Overall, the results suggest that Professor Brownlow's betting profits with the 2:1 bet tend to be positive on average, meaning he loses more money than he wins. The simulation provides insight into the potential range of his profits for different class sizes, and it confirms that the bet is not fair since the average expected profit is negative.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!--> If he consistently makes money (expected profit is positive), he can decrease the odds he offers to balance his expected profits around zero.to find the fair odds (payout ratio), we can set the expected profit to zero and solve for the odds using the BrLw function. Let's update the function to include an odds parameter and then use numerical optimization techniques like the optimize function to find the fair odds.

